Testing Set Features Visualization

### 2022-12-04 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Adding DVF as a tuning removal piece.

Import the file that combines the results from each of the tools from running "combining_tool_output.Rmd":
```{r}
viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
```

@@

```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_tuning_viral=FALSE,
                                include_tuning_not_viral=FALSE,
                                include_virsorter=FALSE) {

@@

  if (include_virsorter2) {
    #keep_score[input_seqs$viral>=50 | input_seqs$lavidaviridae>=0.95 | input_seqs$NCLDV>=0.95] <- keep_score[input_seqs$viral>=50 | input_seqs$lavidaviridae>=0.95 | input_seqs$NCLDV>=0.95] + 0.5
    keep_score[input_seqs$max_score>=50] <- keep_score[input_seqs$max_score>=50] + 0.5
    keep_score[input_seqs$max_score>=95] <- keep_score[input_seqs$max_score>=95] + 0.5
    #keep_score[input_seqs$RNA>=0.95] <- keep_score[input_seqs$RNA>=0.95] + 1
  }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
    keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
  }

  if (include_tuning_viral) {
    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 0.5
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 0.5
    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
  }

  if (include_tuning_not_viral) {
    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] - 1
    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1 
    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
    keep_score[(input_seqs$checkv_completeness<=75 | input_seqs$vibrant_quality=="complete circular")& input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 
  }

@@

@@

  #by type metrics
  fungal_FP <- table(confusion_matrix[seqtype=="fungi"])[2]
  protist_FP <- table(confusion_matrix[seqtype=="protist"])[2]
  bacterial_FP <- table(confusion_matrix[seqtype=="bacteria"])[2]
  viral_FN <- table(confusion_matrix[seqtype=="virus"])[1]

  performance <- c(precision, recall, F1, MCC, auc, fungal_FP, 
                   protist_FP, bacterial_FP, viral_FN)
  names(performance) <- c("precision", "recall", "F1", "MCC", "AUC", "fungal_FP",
                          "protist_FP", "bacterial_FP", "viral_FN")

@@

this function builds a list of all of the combinations that the user wants to 
test. 
In this case, we're comparing the performance of all unique combinations of the 
six tools.
```{r}
build_score_list <- function(input_seqs, combos) {
  output <- data.frame(precision=rep(0, nrow(combos)),
                       recall=rep(0, nrow(combos)),
                       F1=rep(0, nrow(combos)),
                       MCC=rep(0, nrow(combos)),
                       AUC=rep(0, nrow(combos)),
                       fungal_FP=rep(0, nrow(combos)),
                       protist_FP=rep(0, nrow(combos)),
                       bacterial_FP=rep(0, nrow(combos)),
                       viral_FN=rep(0, nrow(combos)))
  for (i in 1:nrow(combos)) {
    keep_score <- getting_viral_set_1(input_seqs, include_vibrant = combos$VIBRANT[i],
                                            include_virsorter = combos$VS[i],
                                            include_virsorter2 = combos$VS2[i],
                                            include_tuning_viral = combos$tune_viral[i],
                                            include_tuning_not_viral = combos$tune_not_viral[i],
                                            include_deepvirfinder = combos$DVF[i])

    output[i,1:9] <- assess_performance(input_seqs$seqtype, keep_score)

@@

## Calculate the performance of each pipeline
```{r}
accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
                      precision=rep(0, nrow(combos_list)*10),
                       recall=rep(0, nrow(combos_list)*10),
                       F1=rep(0, nrow(combos_list)*10),
                       MCC=rep(0, nrow(combos_list)*10), 
                      AUC=rep(0, nrow(combos_list)*10),
                      fungal_FP=rep(0, nrow(combos_list)*10),
                      protist_FP=rep(0, nrow(combos_list)*10),
                      bacterial_FP=rep(0, nrow(combos_list)*10),
                      viral_FN=rep(0, nrow(combos_list)*10))

@@

```{r}
accuracy_scores$numrules <- str_count(accuracy_scores$toolcombo, "1")
#accuracy_scores <- accuracy_scores[order(accuracy_scores$numrules, decreasing=F),]
accuracy_scores <- accuracy_scores[order(accuracy_scores$MCC, decreasing=F),]
accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
accuracy_scores$numrules <- as.factor(accuracy_scores$numrules)
```

@@

## Visualize how the precision, recall, and F1 scores change across pipelines.
```{r}
p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("F1 Score")
p2

ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Precision")
ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Recall")

ggplot(accuracy_scores, aes(x=precision, y=recall, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Precision") +
  ylab("Recall") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 1))

ggplot(accuracy_scores[accuracy_scores$testing_set_index==1,], aes(x=precision, y=recall)) +
  geom_label(alpha=0.7, label=accuracy_scores$toolcombo[accuracy_scores$testing_set_index==1]) +
  geom_point(alpha=0.5, aes(color=numrules, fill=numrules)) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Precision") +
  ylab("Recall")

ggplot(accuracy_scores, aes(x=toolcombo, y=abs(precision-recall), 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Precision-Recall")
ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("MCC") +
  scale_fill_manual(name="Number of Rule Sets",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Rule Sets",
                     values = alpha(rev(pal(6)), 1))

ggplot(accuracy_scores, aes(x=toolcombo, y=AUC, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("AUC")
ggplot(accuracy_scores, aes(x=toolcombo, y=fungal_FP, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Fungal False Positives")

ggplot(accuracy_scores, aes(x=toolcombo, y=protist_FP, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Protist False Positives")

ggplot(accuracy_scores, aes(x=toolcombo, y=bacterial_FP, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Bacterial False Positives")

ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Viral False Negatives")

ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Viral False Negatives")

ggplot(accuracy_scores, aes(x=protist_FP, y=fungal_FP, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Protist FP") +
  ylab("Fungal FP")

ggplot(accuracy_scores, aes(x=recall, y=fungal_FP, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Recall") +
  ylab("Fungal FP")

ggplot(accuracy_scores, aes(x=recall, y=protist_FP, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Recall") +
  ylab("Protist FP")
```

```{r}
accuracy_scores_melt <- accuracy_scores %>% 
  select(testing_set_index, precision, recall, MCC, numrules, toolcombo) %>%
  pivot_longer(cols=c(precision, recall, MCC), 
               names_to="performance_metric",
               values_to="performance_metric_score")
```

@@

>>>>

Import the file that combines the results from each of the tools from running "combining_tool_output.Rmd":
```{r}
viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
```
(this is still the old rules)

@@

# Trying new rule set
this set is based on optimizing for the 
```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_tuning_viral=FALSE,
                                include_tuning_not_viral=FALSE,
                                include_virsorter=FALSE) {

@@

  if (include_virsorter2) {
    keep_score[input_seqs$max_score>=50] <- keep_score[input_seqs$max_score>=50] + 0.5
    keep_score[input_seqs$max_score>=95] <- keep_score[input_seqs$max_score>=95] + 0.5  
  }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5 
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5 
  }

  if (include_deepvirfinder) {
    # remove if DVF calls it not viral
    keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] <- keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] - 1
    # add if DVF calls viral
    keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
    keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
  }

  if (include_tuning_viral) {
  # tuning addition
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
    #note: having kaiju removal rule didn't help
    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
    keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] <- keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] + 0.5 
    keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] <- keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] + 0.5 
  }

  if (include_tuning_not_viral) {
    # tuning removal 
    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- 0
    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1 
    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- 0
    keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] <- keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] - 0.5 
  }

@@

## Trying with all of the rules
```{r}
viruses$keep_score_all_2 <- getting_viral_set_1(viruses,
                                                include_vibrant=TRUE, 
                                                include_virsorter2=TRUE,
                                                include_deepvirfinder=TRUE,
                                                include_tuning_viral=TRUE,
                                                include_tuning_not_viral=TRUE,
                                                include_virsorter=TRUE)

viruses$confusion_matrix_all_2 <- "true negative"
viruses$confusion_matrix_all_2[viruses$seqtype=="virus" & viruses$keep_score_all_2<1] <- "false negative"
viruses$confusion_matrix_all_2[viruses$seqtype=="virus" & viruses$keep_score_all_2>=1] <- "true positive"
viruses$confusion_matrix_all_2[viruses$seqtype!="virus" & viruses$keep_score_all_2>=1] <- "false positive"

viruses$accurate_all_2 <- "false"
viruses$accurate_all_2[grep("true", viruses$confusion_matrix_all_2)] <- "true"
```

accuracy:
```{r}
length(grep("true", viruses$accurate_all_2))/nrow(viruses)
```

precision
```{r}
length(grep("true positive", viruses$confusion_matrix_all_2))/length(grep("positive", viruses$confusion_matrix_all_2))
```

recall
```{r}
length(grep("true positive", viruses$confusion_matrix_all_2))/(length(grep("true positive", viruses$confusion_matrix_all_2))+length(grep("false negative", viruses$confusion_matrix_all_2)))
```

visualizing accuracy by taxa
```{r}
accurate_by_taxa <- melt(table(viruses$accurate_all_2, viruses$seqtype, viruses$Index))
colnames(accurate_by_taxa) <- c("accurate", "seqtype","Index", "count")
```

```{r}
ggplot(accurate_by_taxa, aes(y=count, x=accurate,
                   fill=accurate,
                   color=accurate)) +
  geom_boxplot() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(2)), 0.5),
                    labels=c("false", "true")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(2)), 1),
                    labels=c("false", "true")) +
  xlab("Number of Sequences") +
  ylab("") + 
  facet_wrap(~seqtype, scales = "free")
```

@@

  performance <- c(precision, recall, F1, MCC, auc)
  names(performance) <- c("precision", "recall", "F1", "MCC", "AUC")

@@

this function builds a list of all of the combinations that the user wants to 
test. 
In this case, we're comparing the performance of all unique combinations of the 
six tools.
```{r}
build_score_list <- function(input_seqs, combos) {
  output <- data.frame(precision=rep(0, nrow(combos)),
                       recall=rep(0, nrow(combos)),
                       F1=rep(0, nrow(combos)),
                       MCC=rep(0, nrow(combos)),
                       AUC=rep(0, nrow(combos)))
  for (i in 1:nrow(combos)) {
    keep_score <- getting_viral_set_1(input_seqs, include_vibrant = combos$VIBRANT[i],
                                            include_virsorter = combos$VS[i],
                                            include_virsorter2 = combos$VS2[i],
                                            include_tuning_viral = combos$tune_viral[i],
                                            include_tuning_not_viral = combos$tune_not_viral[i],
                                            include_deepvirfinder = combos$DVF[i])

    output[i,1:5] <- assess_performance(input_seqs$seqtype, keep_score)

@@

## Calculate the performance of each pipeline
```{r}
accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
                      precision=rep(0, nrow(combos_list)*10),
                       recall=rep(0, nrow(combos_list)*10),
                       F1=rep(0, nrow(combos_list)*10),
                       MCC=rep(0, nrow(combos_list)*10), 
                      AUC=rep(0, nrow(combos_list)*10))

@@

```{r}
accuracy_scores$numrules <- str_count(accuracy_scores$toolcombo, "1")
#accuracy_scores <- accuracy_scores[order(accuracy_scores$numrules, decreasing=F),]
accuracy_scores <- accuracy_scores[order(accuracy_scores$MCC, decreasing=F),]
accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
```

accounting for rules that have multiple tools in them
```{r}
accuracy_scores$numtools <- accuracy_scores$numrules

for (i in 1:nrow(accuracy_scores)) {
  toolcombo <- as.character(accuracy_scores$toolcombo[i])
  if (substr(toolcombo, 1, 1)=="1") {
    accuracy_scores$numtools[i] <- accuracy_scores$numtools[i] + 3
  }

  if (substr(toolcombo, 5, 5)=="1") {
    accuracy_scores$numtools[i] <- accuracy_scores$numtools[i] + 4
  }
}

accuracy_scores$numtools[accuracy_scores$numtools>6] <- 6
```

```{r}
accuracy_scores$numrules <- as.factor(accuracy_scores$numrules)
accuracy_scores$numtools <- as.factor(accuracy_scores$numtools)
```

@@

## Visualize how the precision, recall, and F1 scores change across pipelines.
```{r}
p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("F1 Score") +
  scale_fill_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 1))
p2

ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Precision") +
  scale_fill_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 1))

ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Recall") +
  scale_fill_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 1))

ggplot(accuracy_scores, aes(x=precision, y=recall, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Precision") +
  ylab("Recall") +
  scale_fill_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 1))

ggplot(accuracy_scores, aes(x=precision, y=recall, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Precision") +
  ylab("Recall") +
  scale_fill_manual(name="Number of Tools",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Tools",
                     values = alpha(rev(pal(6)), 1))

ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("MCC") +
  scale_fill_manual(name="Number of Rule Sets",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Rule Sets",
                     values = alpha(rev(pal(6)), 1))

ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("MCC") +
  scale_fill_manual(name="Number of Tools",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Tools",
                     values = alpha(rev(pal(6)), 1))
```

```{r}
accuracy_scores_melt <- accuracy_scores %>% 
  select(testing_set_index, precision, recall, MCC, numrules, numtools, toolcombo) %>%
  pivot_longer(cols=c(precision, recall, MCC), 
               names_to="performance_metric",
               values_to="performance_metric_score")
```

@@

ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric=="MCC",], aes(x=numrules, y=performance_metric_score, 
                                  color=numtools, fill=numtools)) +
  geom_boxplot() +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("MCC") +
  xlab("Number of Rule Sets") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 1)) 
```

### 2022-12-04 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
Adding DVF as a tuning removal piece.

@@

```{r}
ggplot(viruses_sankey_tv,
       aes(axis1 = kj_cel, axis2 = hall, axis3 = pv, axis4 = cvl_pu, y=n)) +
  geom_alluvium(aes(fill=viral_score),
                width = 0, knot.pos = 0, reverse = FALSE) +
  geom_stratum(width = 1/5) +
  theme_bw() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            reverse = FALSE) +
  theme(
        axis.text.x=element_text(size=14, angle = 90)
        ) +
  scale_x_continuous(breaks=c(1,2,3,4),
    labels=c("kaiju", "hallmark", "% viral", "% unknown")) +
  facet_wrap(~seqtype, scales="free_y") 
```

>>>>

# Looking at Features

@@

```{r}
ggplot(viruses_sankey_tv,
       aes(axis1 = kj_cel, axis2 = hall, axis3 = pv, axis4 = cvl_pu, y=n)) +
  ggalluvial::geom_alluvium(aes(fill=viral_score),
                width = 0, knot.pos = 0, reverse = FALSE) +
  ggalluvial::geom_stratum(width = 1/5) +
  theme_bw() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            reverse = FALSE) +
  theme(
        axis.text.x=element_text(size=14, angle = 90)
        ) +
  scale_x_continuous(breaks=c(1,2,3,4),
    labels=c("kaiju", "hallmark", "% viral", "% unknown")) +
  facet_wrap(~seqtype, scales="free_y") 
```
